---
title: "Problem set 3: Summarizing distributions"
author: "(Your name here)"
subtitle: "Due October 21, 2024, at 10am"
format: pdf
---


<!-- Define special characters for operators -->
\def\E{{\textrm E}\,}
\def\V{{\textrm V}\,}



*\textsc{Note}: Start with the file `ps3_2024_summarizing_distributions.qmd` (available from the github repository at <https://github.com/UChicago-pol-methods/IntroQSS-F24/tree/main/assignments>). Modify that file to include your answers. Make sure you can "render" the file (e.g. in RStudio by clicking on the `Render` button). Submit both the qmd file and the PDF via Canvas.* 




## Question 1: Expected value 

Consider the random variable $X$ characterized by the following PMF.

| $x$  | $f(x)$ |
|:------:|:------------:|
|   0   |     .3     |
|   1   |     .5     |
|   2   |     .15     |
|  3 |     .05      |

\vspace{.1in}

(1a) Compute $\E[X]$. Show your work.

**Answer**:

$$
\begin{aligned} \E[X] &= \sum_x x f(x) \\
  &= 0 \times .3 + 1 \times .5 + 2 \times .15 + 3 \times .05 \\
  &= .95 \end{aligned}  
$$
\vspace{.1in}
 
 
(1b) Write an `R` function to compute the expectation of any discrete random variable. The arguments to your function should include the values the random variable can take on (`x`) and the probability it takes on each value (`probs`). Use your function to confirm your answer from question (1a).

**Answer**:

```{r}
exp_func <- function(x, probs){
  sum(x * probs)
}
x_vals <- c(0, 1, 2, 3)
x_probs <- c(.3, .5, .15, .05)
exp_func(x = x_vals, probs = x_probs)
```

\vspace{.1in}


(1c) Compute the MSE ($\E[(X - c)^2]$) for $c = 0$ and $c = 1$. Show your work.

**Answer**:

$$
\begin{aligned} \E[(X - 0)^2] &= \sum_x (x - 0)^2 f(x) \\
  &= (0 - 0)^2 \times .3 + (1 - 0)^2 \times .5 + (2 - 0)^2 \times .15 + (3 - 0)^2 \times .05 \\
  &= 0 + .5 + .6 + .45 \\ 
  &= 1.55 \end{aligned}  
$$

$$
\begin{aligned} \E[(X - 1)^2] &= \sum_x (x - 1)^2 f(x) \\
  &= (0 - 1)^2 \times .3 + (1 - 1)^2 \times .5 + (2 - 1)^2 \times .15 + (3 - 1)^2 \times .05 \\
  &= .3 + 0 + .15 + .2 \\ 
  &= .65 \end{aligned}  
$$

\vspace{.1in}


(1d) Write a function to compute the MSE for any discrete random variable at a value `c`. The arguments to your function should include the values the random variable can take on (`x`), the probability it takes on each value (`probs`), and the value `c` being considered. Use your function to confirm your answers from (1c).

**Answer**:

```{r}
mse_func <- function(x, probs, c){
  sum(probs*(x - c)^2)
}
mse_func(x = x_vals, probs = x_probs, c = 0)
mse_func(x = x_vals, probs = x_probs, c = 1)
```

\vspace{.1in}


(1e) Create a vector `cs` that contains numbers in the sequence `0.5, 0.55, 0.6, 0.65, 0.7, . . ., 2`. Using a for-loop and your function from (1d), compute the MSE for the random variable $X$ whose PMF was given above at each value of `cs` and store the result in a vector called `mses`.

**Answer**:

```{r}
cs <- seq(from = .5, to = 2, by = .05)
mses <- rep(NA, times = length(cs))
for(i in 1:length(mses)){
  mses[i] <- mse_func(x = x_vals, probs = x_probs, c = cs[i])
}
```

\vspace{.1in}


(1f) Using the `plot()` command, make a scatterplot showing the MSE of $X$ (vertical axis) at each value of $c \in \{0.5, 0.55, 0.6, \ldots, 2\}$ (horizontal axis). Use `abline()` to add a vertical line at $E[X]$.

**Answer**:

```{r}
plot(cs, mses)
abline(v = .95)
```

\vspace{.2in}

## Question 2: Variance 

Consider the random variable $X$ characterized by the following PMF:

| $x$   | $f(x)$  |
|:-----:|:----------:|
|   1   |     .2     |
|   2   |     .7     |
|   3   |     .1     |

\vspace{.1in}

(2a) Confirm that the variance of $X$ is the same whether we compute it by the formula in Definition 2.1.12 or the Alternative Formula in Theorem 2.1.13. (We want the two variance computations, not the proof.)


**Answer**:
First we calculate $\E[X]$: 

$$
\begin{aligned}
  \E[X] &= \sum_x x f(x)\\
  & = 1 \times .2 + 2 \times .7 + 3 \times .1 \\
  & = 1.9
\end{aligned}
$$

Using Definition 2.1.12:
$$
\begin{aligned} 
  \V[X] & = \E\left[(X - \E[X])^2\right] \\
  &=  \E\left[(X - 1.9)^2\right]\\
  &=  (-.9)^2 \times .2 + .1^2 \times .7 + 1.1^2 \times .1 \\
  &= .29  \end{aligned}  
$$

To use Theorem 2.1.13, we calculate $\E[X^2]$: 

$$
\begin{aligned}
  \E[X^2] &= \sum_x x^2 f(x)\\
  & = 1 \times .2 + 4 \times .7 + 9 \times .1 \\
  & = 3.9
\end{aligned}
$$

Using Theorem 2.1.13:
$$
\begin{aligned} 
  \V[X] & = \E[X^2]  - \E[X]^2 \\
  &=  3.9 - 3.61\\
  &=  .29 \end{aligned}  
$$
\vspace{.1in}

(2b) Write an `R` function to compute the variance of any discrete random variable. The arguments to your function should include the values the random variable can take on (`x`) and the probability it takes on each value (`probs`). Use your function to confirm your answer from question (2a).

**Answer**: 


```{r}
# two versions, both of which use the expectation function I wrote above 

# this version uses the E[(X - E[X])^2] formulation
var_func2 <- function(x, probs){
  ex <- exp_func(x, probs)
  exp_func((x - ex)^2, probs)
}

# this version uses the E[X^2] - E[X]^2 formulation
var_func <- function(x, probs){
  exp_func(x^2, probs) - exp_func(x, probs)^2
}

# confirming equality
var_func(x = c(1, 2, 3), probs = c(.2, .7, .1))
var_func2(x = c(1, 2, 3), probs = c(.2, .7, .1))
```

\vspace{.1in}

(2c) Explain what definition/property/mathematical operation is being used in each step of the following proof.

For random variable $X$ and $a \in \mathbb{R}$, $\V[aX] = a^2\V[X]$,

\begin{align}
\V[aX] &= \E\left[(aX - \E[aX])^2\right]  \tag{Step 1} \\
&= \E\left[(aX - a\E[X])^2\right]  \tag{Step 2} \\
&= \E\left[a^2(X - \E[X])^2\right]  \tag{Step 3} \\
&= a^2 \E\left[(X - \E[X])^2\right]  \tag{Step 4} \\
&= a^2 \V[X] \tag{Step 5}
\end{align}


**Answer**: 

In Step 1, . . .

In Step 2, . . . 

In Step 3, . . . 

In Step 4, . . . 

In Step 5, . . . 


**Answer**: 

In Step 1, we use the **definition of variance** applied to the random variable $aX$.  

In Step 2, we move a constant outside of the expectations operator using **the property $E[aX] = aE[X]$ that is stated in Theorem 2.1.6**, and that also follows from Linearity of Expectations. This property follows from the fact that (with discrete $X$) an expectation is a (weighted) sum, and in this case each element of that sum is multiplied by $a$, so it can be brought outside the summation. 

In Step 3, we do a mathematical operation to **move $a$ outside of the parentheses**. In more detail, we first express $aX - a\E[X]$ as $a(X - \E[X])$, then we apply the square to both of them, i.e. $\left(a(X - \E[X])\right)^2 = a^2(X - \E[X])^2$.

In Step 4, again move a constant outside of the expectations operator using **the property $E[aX] = aE[X]$** (see Step 2).

In Step 5, we use the **definition of variance**. 

\vspace{.1in}

(2d) Use your variance function to confirm that $\V[aX] = a^2\V[X] \forall a \in \mathbb{R}$. (You can just show this for one value of $a$.)

**Answer**

```{r}
a <- 5
var_func(x = a*c(0, 1, 2), probs = c(.6, .1, .3))
var_func(x = c(0, 1, 2), probs = c(.6, .1, .3))*a^2
```

\vspace{.1in}

(2e) Use your function to confirm that $V[X + c] = V[X] \, \forall \, c \in \mathbb{R}$. (You can just show this for one value of $c$.)

**Answer**

```{r}
c <- 6
var_func(x = c(0, 1, 2), probs = c(.6, .1, .3))
var_func(x = c(0, 1, 2) + 6, probs = c(.6, .1, .3))
```




