---
title: "Midterm Solutions"
format: 
  pdf:
    include-in-header:
      text: '\pagenumbering{gobble}'
author: "Name: _______________________"
---

<!-- Define special characters for operators -->
\def\E{{\textrm E}\,}
\def\V{{\textrm V}\,}


**Question 1: `R` code interpretation**

Consider this `R` code.

```{r}
#| eval: true
artist <- c("da vinci", "da vinci", "caravaggio", "delacroix", "da vinci")
area <- c(5, 6, 9, 12, 8)
rating <- c(7, 6, 10, 2, 5)
```

(1a) What is `mean(rating)`? 
$$
\frac{7+6+10+2+5}{5} = 6
$$

```{r}
#check
mean(rating)
```


\vspace{0.8in}

(1b) What is `mean(rating[artist == "da vinci"])`?
$$
\frac{7+6+5}{3} = 6
$$

```{r}
#check
mean(rating[artist == "da vinci"])
```

\vspace{0.8in}

(1c) What is `mean(rating[area > 8])`?
$$
\frac{10+2}{2} = 6
$$

```{r}
#Check
mean(rating[area > 8])
```


\vspace{0.5in}

```{r}
# From R help file: 
# `sample` takes a sample of the specified `size` from the elements of `x` 
# using either with or without replacement.
samp <- sample(x = area, size = 20000, replace = T, prob = rep(1/5, 5))
```

(1d) What is `mean(samp)`, approximately?
$$
\frac{5+6+9+12+8}{5}=8
$$
Because the sampling is done with replacement, and because each element in the area vector is equally likely to be selected, `mean(samp)` should be approximately 8. 
```{r}
#check
mean(samp)
```

\vspace{0.8in}

\newpage

**Question 2: Proof interpretation**

The variance of a random variable $X$ is defined as  $\V[X] = \E\left[\left(X - \E[X]\right)^2\right]$. An alternative formula for variance is 
$$\V[X] = \E[X^2] - \E[X]^2.$$
Here is a proof:
\begin{align}
\V[X] &= \E\left[\left(X - \E[X]\right)^2\right] \tag{Step 1} \\
   &= \E\left[X^2 - 2\E[X] X  + \E[X]^2 \right] \tag{Step 2} \\ 
   &= \E[X^2] - \E\left[2\E[X] X \right] + \E\left[ \E[X]^2 \right] \tag{Step 3} \\
   &= \E[X^2] - 2\E[X] \E[X] + \E\left[ \E[X]^2 \right] \tag{Step 4} \\
   &= \E[X^2] - 2\E[X] \E[X] + \E[X]^2 \tag{Step 5} \\
   &= \E[X^2] - \E[X]^2 \tag{Step 6} 
\end{align}

Explain what definition/property/mathematical operation is being used in each step of the proof.

(2a) Step 1: Definition of Variance. 

\vspace{.6in}

(2b) Step 2: Expand the squared term. 

\vspace{.6in}

(2c) Step 3: Linearity of Expectations. 

\vspace{.6in}

(2d) Step 4: Linearity of Expectations (moving $2\E[X]$ outside of the expectations operator, because it is a constant)

\vspace{.6in}

(2e) Step 5: Expectation of a constant is a constant. (Could also say linearity of expectations, or moving a constant outside of the expectations operator.)

\vspace{.6in}

(2f) Step 6: Combine like terms. 

\vspace{.6in}



**Question 3: Joint distribution of two random variables**

<!-- is this question already okay? --> 

Consider the joint PMF of two random variables, $X$ and $Y$:

| $x$   | $y$ | $f(x, y)$    |
|:-----:|:-----:|:----------:|
|   0   |   0    |    1/6    |
|   1   |   0    |    1/3    |
|   2   |   0    |    1/8    |
|   2   |    1   |   3/8     |


(3a) What is the marginal distribution of $X$, i.e. $f_X(x)$?

| $x$   | $f_{X}(x)$  |
|:-----:|:-----:|
|   0   |  1/6  |       
|   1   |  1/3  |       
|   2   |  1/2  | 

(3b) What is the expectation of $X$, i.e. $\E[X]$? 
$$
\E[X] = 0*\frac{1}{6}+1*\frac{1}{3}+2*\frac{1}{2}=\frac{4}{3}
$$

\vspace{1in}

(3c) What is the variance of $X$, i.e. $\V[X]$? 

Given $\V[X] = \E[X^2] - \E[X]^2$, 
\begin{align}
\V[X] &= \left(0^2*\frac{1}{6}+1^2*\frac{1}{3}+2^2*\frac{1}{2}\right) - \left(\frac{4}{3}\right)^2 \\
&= \frac{7}{3} - \frac{16}{9} \\
&= \frac{5}{9}
\end{align}



\vspace{1in}

(3d) What is the conditional distribution of $Y$ given $X$, i.e. $f_{Y\mid X}(y \mid x)$?  


| $x$   | $y$ | $f_{Y\mid X}(y \mid x)$  |
|:-----:|:-----:|:----------:|
|   0   |   0    |    1   |
|   1   |   0    |    1   |
|   2   |   0    |   1/4  | 
|   2   |   1    |   3/4  |

